Algebraic File Synchronization: Adequacy and Completeness

### 2016-10-08 p2/def-commands.tex
Fixed lemma and proof

\begin{mydef}[Regular commands]
Therefore we have three pieces of information related to each command,
and represent them using triples that contain
their input type ($[v_{in}]$), output value ($v_{out}$),
and the node to which they are applied ($n$):
\[ \caaaa{v_{in}}{v_{out}}{n}. \]
\end{mydef}

>>>>

\begin{mydef}[Regular commands]
Therefore we have three pieces of information related to each command,
and represent them using triples that contain
their input type ($[v_\textrm{in}]$), output value ($v_\textrm{out}$),
and the node to which they are applied ($n$):
\[ \caaaa{v_\textrm{in}}{v_\textrm{out}}{n}. \]
\end{mydef}

### 2016-10-08 p2/works-inverse.tex
Fixed lemma and proof

We select representatives of each $\typeeq$-equivalence class of values:
for $[\empt]$ and $[\vald]$ this must be $\empt$ and $\vald$; and
for $[\valf]$ we select an arbitrary file value $\fil\in[\valf]$.
We then use these representative values to
select representatives of the equivalence classes of
other objects in the following way:

\begin{mydef}[$\T$]
We use $\T$ and write $X\T$ to denote the canonical mapping of 
values, filesystems, commands and sequences to the
representatives of their $\typeeq$-equivalence class.
\begin{itemize}
\item For values, $\vald\T=\vald$, $\valf\T=\fil$ and $\empt\T=\empt$.
\item For filesystems, $\FS\T$ is defined by $\FS\T(n) = \FS(n)\T$.
\item For commands, $\cxynv\T = \cTxynv$.
\item For sequences of commands, $\emptyseq\T=\emptyseq$, and if $S'=\alpha\cc S$, then $S'\T=\alpha\T\cc S\T$.
\item Finally we extend $\T$ to sets of sequences: $\seqset{A}\T = \{S\T | S\in\seqset{A}\}$.
\end{itemize}
\end{mydef}

It is easy to see that for all objects where it is defined, $X\T\typeeq X$.

In other words, $\T$ maps our system of filesystems and commands
into a \emph{representative} system where there is a one-to-one correspondence between types
and values.
% $|\setf|$ is also $1$.
At the same time, representative filesystems, commands, sequences, etc.
are subsets of their non-representative counterparts; hence all
propositions proven so far apply to them as well.

% TODO <[x],y,n> T = <[x],[y],n>
% one-to-one corresp btw values and types

\begin{mycor}
For any filesystem $\FS$ and command $\alpha$, 
$\alpha\T(\FS\T) = (\alpha\aFS)\T$.
\end{mycor}

@@

As whether a sequence of commands breaks a filesystem does not depend on the output values
of the commands, only the input and output types, we know that
\begin{mycor}
\begin{gather*}
\works{\seqset{A}} \Leftrightarrow \works{\seqset{A}\T} \\
\worksc{\seqset{A}}{\seqset{B}} \Leftrightarrow \worksc{\seqset{A}\T}{\seqset{B}\T}.
\end{gather*}
\end{mycor}

@@

\begin{mydef}[Inverse commands and sequences]
If based on \cref{repr_comm_inject}
we regard the representative command $\alpha\T$ a bijective function whose
domain is the subset of representative filesystems $\{\FS\T | \alpha\aFS\neq\fsbroken\}$,
then we can define the inverse of $\alpha\T$
for which we write $\alpha\T^{-1}$.
For any $\GS\T$ outside the range of $\alpha\T$ we define $\alpha\T^{-1}(\GS\T)$ to be $\fsbroken$.

@@

%% TODO
% Do we need <[x],[y],n>-1 = <[y],[x],n> ?
% Do we need to prove if a filesystem is not in the image, then a-1 will break it?
\begin{mylem}
For representative commands
\[ \cTxynv^{-1} = \cTyxnv \]
\end{mylem}
TODO

@@

\begin{mylem}\label{r_invmove}
A common initial segment of a set of sequences can be moved to the other side of $\workssign$ by inverting it:
\begin{gather*}
\worksc{B\cc \seqset{A}}{\seqset{C}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{C}} \\
\worksc{\seqset{A}}{B\cc \seqset{C}} \Rightarrow \worksc{B^{-1}\cc \seqset{A}}{\seqset{C}}
\end{gather*}
\end{mylem}
\begin{proof}

%% TODO see lemma above

This is based on the fact that in our model, unless they break a filesystem,
commands leave the type of the filesystem value unchanged 
or change one type into another, but never merge types.
That is,
sequences---as functions mapping filesystems to filesystems---are
essentially bijections over type-equality ($\typeeq$)
with the only ``sink'' being $\fsbroken$:
\begin{align*}
\forall S,\FS,\GS:{}&S\aFS\neq\fsbroken \Rightarrow \\
&(S\aFS\ntypeeq S\aGS \Longleftrightarrow \FS\ntypeeq \GS).
\end{align*}

For the proposition
\[ \worksc{B\cc \seqset{A}}{\seqset{C}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{C}} \]
the rest of the proof is illustrated by \cref{fig_invmove}
where $\textrm{Dom}(x)$, the domain of $x$, represents the set of filesystems
the sequence $x$ does not break, or neither of the sequences in the set $x$ breaks.

@@

We see that the set of filesystems the sequences in $\seqset{A}$ do not break intersect with the range of $B$.
As $B$ is a bijection between its domain and range, we can use $B^{-1}$ to map this intersection back
onto the domain of $B$.
As $\worksc{B\cc \seqset{A}}{\seqset{C}}$ the domain of $\seqset{C}$ must be a subset of this
projected intersection.
If so, then we can use $B$ to map the domain of $\seqset{C}$, which yields the domain of $B^{-1}\cc \seqset{C}$.
As it is also a part of the domain of $A$, we get $\worksc{\seqset{A}}{B^{-1}\cc \seqset{C}}$.

The second part of the lemma can be proven in a similar way.
\end{proof}

\begin{mylem}\label{indep_prefix_combine}
The combination of sequences with a common head and independent tails 
continues to work under the same conditions:
\[ \worksc{A\cc B}{\seqset{S}} \wedge \worksc{A\cc C}{\seqset{S}} \wedge B\indep C \Rightarrow \worksc{A\cc B\cc C}{\seqset{S}} \]
\end{mylem}
\begin{proof}
Based on \cref{r_invmove} we know
$\worksc{B}{A^{-1}\cc \seqset{S}}$ and $\worksc{C}{A^{-1}\cc \seqset{S}}$,
and from \cref{combine_independent_sequences} we know
$\worksc{B\cc C}{B,C}$.
Combining these using \cref{workschained}
we get
$\worksc{B\cc C}{A^{-1}\cc \seqset{S}}$. 
Finally, applying the second line of \cref{r_invmove} yields
$\worksc{(A^{-1})^{-1}\cc B\cc C}{\seqset{S}}$, which proves our lemma 
as $\forall\FS: (A^{-1})^{-1}\aFS\typeeq A\aFS$ (\cref{negneg_is_typeeq}).
\end{proof}

>>>>

We select representatives of each $\typeeq$-equivalence class of values:
for $[\empt]$ and $[\vald]$ this must be $\empt$ and $\vald$; and
for $[\valf]$ we select an arbitrary file value $\fil\in[\valf]$.
We then use these representative values to
introduce the canonical mapping $\T$ that preserves the types of values,
but maps all objects into a subsystem where there is a one-to-one
correspondence between types and values.

\begin{mydef}[$\T$]
We use $\T$ and write $X\T$ to denote the canonical mapping of 
values, filesystems, commands and sequences to the
representatives of their $\typeeq$-equivalence class.
\begin{itemize}
\item For values, $\vald\T=\vald$, $\valf\T=\fil$ and $\empt\T=\empt$.
\item For filesystems, $\fsbroken\T=\fsbroken$ and in other cases $\FS\T$ is defined by $\FS\T(n) = \FS(n)\T$.
\item For commands, $\cbrk\T=\cbrk$ and $\cxynv\T = \cTxynv$.
\item For sequences of commands, $\emptyseq\T=\emptyseq$, and if $S'=\alpha\cc S$, then $S'\T=\alpha\T\cc S\T$.
\item Finally we extend $\T$ to sets of sequences: $\seqset{A}\T = \{S\T | S\in\seqset{A}\}$.
\end{itemize}
\end{mydef}

It is easy to see that for all objects where it is defined, $X\T\typeeq X$.
% $|\setf|$ is also $1$.
As representatives of filesystems, commands, sequences, etc.
are subsets of their non-representative counterparts, all
propositions proven so far apply to them as well.

\begin{mycor}
For any filesystem $\FS$ and command $\alpha$, 
$\alpha\T(\FS\T) = (\alpha\aFS)\T$.
TODO needed?
\end{mycor}

@@

As whether a sequence of commands breaks a filesystem does not depend on the output values
of the commands, only the input and output types, we know that
\begin{mycor}\label{repr_works_is_same}
\begin{gather*}
\works{\seqset{A}} \Leftrightarrow \works{\seqset{A}\T} \\
\worksc{\seqset{A}}{\seqset{B}} \Leftrightarrow \worksc{\seqset{A}\T}{\seqset{B}\T}.
\end{gather*}
\end{mycor}

@@

\begin{mydef}[Inverse commands and sequences]
If based on \cref{repr_comm_inject}
we regard the representative command $\alpha\T$ a bijective function whose
domain is the subset of representative filesystems $\{\FS\T | \alpha\aFS\neq\fsbroken\}$,
then we can define the inverse of $\alpha\T$
for which we write $\alpha\T^{-1}$.
For any $\GS\T$ outside the image of the domain of $\alpha\T$ we define $\alpha\T^{-1}(\GS\T)$ to be $\fsbroken$.

@@

%% TODO
% Do we need <[x],[y],n>-1 = <[y],[x],n> ?
% Do we need to prove if a filesystem is not in the image, then a-1 will break it?
\begin{mylem}
For representative commands
\[ \cTxynv^{-1} = \cTyxnv. \]
\end{mylem}
TODO needed? -maybe for \cref{negneg_is_typeeq}

@@

For brevity, in the following lemma we use $B$, $B'$, $\seqset{A}$ and $\seqset{C}$ to denote
representative sequences and sets.

\begin{mylem}\label{r_invmove}
A common initial segment of a set of \emph{representative} sequences can be moved to the other side of $\workssign$ by inverting it:
\begin{gather*}
\worksc{B\cc \seqset{A}}{\seqset{C}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{C}} \\
\worksc{B}{\seqset{C}} \wedge \worksc{\seqset{A}}{B^{-1}\cc \seqset{C}} \Rightarrow \worksc{B\cc \seqset{A}}{\seqset{C}}
\end{gather*}
\end{mylem}
\begin{proof}

The proof for both propositions is illustrated by \cref{fig_invmove}, where
$\Dom{X}$, the domain of $X$, represents the set of filesystems
the sequence $X$ does not break, or neither of the sequences in the set $X$ breaks.
We write $\Img{X}$ to denote the image of the domain of the sequence $X$.
We also note that in general $\worksc{\seqset{X}}{\seqset{Y}} \Leftrightarrow \Dom{\seqset{Y}} \subset \Dom{\seqset{X}}$.

@@

Unless $\works{\seqset{C}}$, both propositions are trivially true. Otherwise,
for the first proposition
we know $\works{B\cc\seqset{A}}$, which implies that $\Dom{\seqset{A}}$ intersects with $\Img{B}$ 
as otherwise $B\cc\seqset{A}$ would break all filesystems.
As the (representative) $B$ is a bijection between its domain and image,
we can use $B^{-1}$ to map this intersection back
onto the domain of $B$ and get $\Dom{B\cc\seqset{A}}$.
As $\worksc{B\cc \seqset{A}}{\seqset{C}}$, 
we know that $\Dom{\seqset{C}} \subset \Dom{B\cc\seqset{A}}$.
We now use $B$ to map both sides of this statement, which yields
$\Dom{B^{-1}\cc \seqset{C}} \subset \Dom{B^{-1}\cc B\cc\seqset{A}}$,
but the latter is a subset of $\Dom{\seqset{A}}$,
and so $\Dom{B^{-1}\cc \seqset{C}} \subset \Dom{\seqset{A}}$,
from which $\worksc{\seqset{A}}{B^{-1}\cc \seqset{C}}$.

For the second proposition, 
from $\worksc{B}{\seqset{C}}$ we know $\Dom{\seqset{C}} \subset \Dom{B}$.
Therefore we can project the whole of $\Dom{\seqset{C}}$ using the bijective representative sequence $B$
to get $\Dom{B^{-1}\cc\seqset{C}}$, and as $\worksc{\seqset{A}}{B^{-1}\cc \seqset{C}}$
we know
$\Dom{B^{-1}\cc\seqset{C}} \subset \Dom{\seqset{A}}$.
We use $B^{-1}$ to project both sides of this statement to get
$\Dom{B\cc B^{-1}\cc\seqset{C}} \subset \Dom{B\cc\seqset{A}}$.
But we also know that $\Dom{B\cc B^{-1}\cc\seqset{C}} = \Dom{\seqset{C}}$
as $\Dom{\seqset{C}} \subset \Dom{B}$,
and so $\Dom{\seqset{C}} \subset \Dom{B\cc\seqset{A}}$
from which $\worksc{B\cc \seqset{A}}{\seqset{C}}$.
\end{proof}

\begin{mylem}\label{indep_prefix_combine}
The combination of sequences with a common head and independent tails 
continues to work under the same conditions:
\[ \worksc{A\cc B}{\seqset{S}} \wedge \worksc{A\cc C}{\seqset{S}} \wedge B\indep C \Rightarrow \worksc{A\cc B\cc C}{\seqset{S}} \]
\end{mylem}
\begin{proof}
Based on \cref{repr_works_is_same} we have
$\worksc{A\T\cc B\T}{\seqset{S\T}}$ and $\worksc{A\T\cc C\T}{\seqset{S\T}}$,
and from \cref{r_invmove} we know
$\worksc{B\T}{{A\T}^{-1}\cc \seqset{S\T}}$ and $\worksc{C\T}{{A\T}^{-1}\cc \seqset{S\T}}$.
Using \cref{combine_independent_sequences} we get
$\worksc{B\T\cc C\T}{B\T,C\T}$,
and combining these using \cref{workschained}
we get
$\worksc{B\T\cc C\T}{{A\T}^{-1}\cc \seqset{S\T}}$.

From $\worksc{A\cc B}{\seqset{S}}$ we also know $\worksc{A}{\seqset{S}}$ and $\worksc{A\T}{\seqset{S\T}}$.
We can combine this and $\worksc{B\T\cc C\T}{{A\T}^{-1}\cc \seqset{S\T}}$ using
the second proposition of \cref{r_invmove} to get
$\worksc{A\T\cc B\T\cc C\T}{\seqset{S\T}}$, and from \cref{repr_works_is_same} we know
$\worksc{A\cc B\cc C}{\seqset{S}}$.
\end{proof}
